### 2024-10-08 Docs/linear-reduction.tex
Worked on Makefiles, README, comparisons
Updated documentation

\input pdfToolbox
\input preamble

@@

In this paper we will define the concept of linear reduction in the context of syntax parsing.
We will progress through more and more complicated examples, beginning from the programming of a simple calculator until we ultimately have created an extensible programming language, which we call Lang-X.

@@

In this section we discuss the grammar of the language of Lang-X.
This is not naturally imposed by the parser, but it will properly parse programs of this form.

@@

We now describe the initial beta reducer of Lang-X according to stateful reduction.
By convention, {\astyle atomic abstract terms} will be red, {\tstyle type terms} will be green, {\istyle internal terms} will be blue.

@@

>>>>

\input pdfToolbox
\input syntax-hi
\input preamble

@@

In this paper we will define the concept of linear reduction in the context of syntax parsing.
We will progress through more and more complicated examples, beginning from the programming of a simple calculator until we ultimately have created an extensible programming language, which we call L-Lang.

@@

In this section we discuss the grammar of the language of L-Lang.
This is not naturally imposed by the parser, but it will properly parse programs of this form.

@@

We now describe the initial beta reducer of L-Lang according to stateful reduction.
By convention, {\astyle atomic abstract terms} will be red, {\tstyle type terms} will be green, {\istyle internal terms} will be blue.

@@

\section{Comparing Runtimes}

For this paper, we developed an interpreter using both our algorithm and the tool menhir.
We then compare runtimes for various different programs.

\subsection{Currying}

We begin with the following program:

\beginhi
fun print (x) {
    _prim_print x
}

fun curry (f) {
    fun curried (x) {
        fun curriedX (y) {
            f (x,y)
        }
        curriedX
    }
    curried
}

fun plus (x,y) {
    x + y
}

print (plus (10, 20));
let curry_plus = curry plus;
print ((curry_plus 10) 20);
\endhi

We compare the time it takes our algorithm and the classical algorithm to run this program 300 times.

\medskip
\pdfximage width .5\hsize{Images/currying.png}
\centerline{\pdfrefximage\the\pdflastximage}
\medskip

It takes our algorithm 0.0040886 seconds to run this program on average, and it takes the classical algorithm 0.0036926 seconds.

\subsection{Expressions}

We compare runtimes of programs of the form: {\tt let x = <expr>;}, where {\tt<expr>} is an arithmetical expression generated by the following algorithm:

\medskip
\algorithm
\Comment Create an expression with $\ell$ subterms, with a parentheses depth of $d$, where numbers are chosen from\cr $[0,n)$.\EndComment
\Function{Generate-Expression}{$\ell,d,n$}
    \lIf{$d=0$} \vtop{\hsize=.8\hsize
        \noindent \Return an expression with $\ell$ random numbers in $[0,n)$ and $\ell-1$ random operators in $\set{+,-,\times,\div}$.}
    \State \Return $\bigl(\hbox{\tencsc Generate-Expression}(\ell,d-1,n)\bigr)\circ_1\cdots\circ_{\ell-1}\bigl(\hbox{\tencsc Generate-Expression}(\ell,d-1,n)\bigr)$
    \nonum\State where $\circ_i\in\set{+,-,\times,\div}$ are chosen randomly
\EndFunc
\ealgorithm
\medskip

Running this with $n=100$, and we parameterize $\ell=d$, we get the following:

\medskip
\pdfximage width .5\hsize{Images/expr.png}
\centerline{\pdfrefximage\the\pdflastximage}
\medskip

\subsection{Fibonacci}

We compare runtimes of a fibonacci program, which computes the $n$th fibonacci number using its recursive formula, where $n$ varies.

\medskip
\pdfximage width .5\hsize{Images/fib.png}
\centerline{\pdfrefximage\the\pdflastximage}
\medskip

Using exponential regression, we estimate that our algorithm takes $ab^n$ seconds, where $a=0.000138585$ and $b=1.64685$; and the classical algorithm takes $ab^n$ seconds where $a=0.00000362553$ and
$b=1.5809$.
